---
title: "Simulation study exploring exposure distribution changes"
output:
  html_document:
    toc: true
    number_sections: true
    theme: cosmo
    toc_depth: 2
---

```{r lib, warning = F, message = F, echo = F}
library(dplyr)
library(reshape2)
library(knitr)
library(ggplot2)
library(RColorBrewer)
library(fda)
library(fda.usc)
library(truncnorm)

opts_chunk$set(message = F, fig.align = "center", warning = F, echo = F)

```

Suppose our standard day of personal exposures looks as below:

```{r, fig.cap = "Histogram of baseline low exposures"}
N <- 1000
x1 <- rlnorm(N)
hist(x1, xlab = "PM", main = "")
```


# Possible distribution changes in exposures



```{r}
plotx <- function(x1, x2) {
  
  xdat <- data.frame(x1, x2)
  mx <- melt(xdat)
  labs1 <- c("Low exposures", "High exposures")
  mx$variable <- factor(mx$variable, labels = labs1)
  cols1 <- c("lightsteelblue", "darksalmon")
  ggplot(mx, aes(x = value, fill = variable)) + geom_density(alpha = 0.3) + 
    scale_fill_manual(name = "", values = cols1) + xlab("PM exposure") + theme_bw()  
  
  
}


plotq <- function(x1, x2, na) {
  argvals1 <- seq(0.05, 0.95, length = na) 

  q1 <- quantile(x1, probs = argvals1)
  q2 <- quantile(x2, probs = argvals1)

  mq <- melt(data.frame(q1, q2, argvals1), id.vars = "argvals1")

  
  cols1 <- c("lightsteelblue", "darksalmon")
  
  labs1 <- c("Low exposures", "High exposures")
  mq$variable <- factor(mq$variable, labels = labs1)

  ggplot(mq, aes(x = argvals1, y = value, color = variable)) + geom_point() +
  
    scale_color_manual(name = "", values = cols1) + theme_bw() +  
  
    xlab("Quantile") + ylab("Concentration")
}
```

## Exposure shift, where all people have the same increase in exposure"

```{r}

x2 <- x1 + 1
plotx(x1, x2)
plotq(x1, x2, 20)

```



## Changes in highest exposure

This could occur on a bad weather day if people who spent time outdoors (e.g. for work) get the brunt of the exposure.  People indoors don't have appreciable changes in exposure.


```{r}

x2 <- x1 + x1 * pnorm(x1, mean = 5, sd = 2)

plotx(x1, x2)
plotq(x1, x2, 20)
```


## Changes in the lowest exposures

This could occur on a nice day, when pollution isn't higher, but the people who don't normally go outdoors spend time outside.

```{r}


x2 <- x1 + 2 * (1 - pnorm(x1, mean = 0.1, sd = 0.5))

plotx(x1, x2)
plotq(x1, x2, 20)

```



## Larger variance in exposures (except right tail)

This could occur on nice days when more people try and go outside.

```{r}

x2 <- rtruncnorm(1000, a = 0, sd = 5, mean = 1)


plotx(x1, x2)
plotq(x1, x2, 20)
```



# Different coefficient functions

## Constant coefficient function

This has the interpretation of the effect of pollution on outcome being the same for all parts of the distribution

```{r}
N <- 1000
seq0 <- seq(0, 1, length = N)
plot1 <- function(x, seq1 = seq0) {

  maxy <- max(x)
  plot(seq1, x, xlim = c(0, 1), ylim = c(0, maxy), type = "l", 
    ylab = "Beta function", xlab = "Quantile of exposure")
  abline(h = 0, lty = 2)
}

plot1(rep(1, N))
```



## Coefficient function increases for highest exposures


```{r}
x1 <- 1 + exp(seq0 * 8) / 1000
plot1(x1)
```


## Coefficient function increases for lowest exposures

```{r}
x1 <- 1 + exp(seq0 * -10) / .5
plot1(x1)

```


## Coefficient function increases for lowest and highest exposures

```{r}
x1 <- 5 * (seq0 - 0.5)^2 + 1
plot1(x1)

```
